### 2015-05-04 docs/parasite_mortatlity_manuscript.tex
Incorporated most of Sara's comments. Still need citations

% Document
\documentclass[12pt, a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage{lipsum}

% Fig. and table formating
\usepackage{epsfig}
\usepackage{tabu}
\usepackage{rotating}
\usepackage{pbox}
\usepackage{framed, multicol}
\usepackage[framemethod=TikZ]{mdframed}
\usepackage{float}
\usepackage[left=1.25 in, right=1.25 in, top=1.25 in, bottom=1.25 in]{geometry}

\usepackage{caption}

% Fonts - Mathtime
%\usepackage{txfonts}
\usepackage{amsmath} % Add amssymb if not using Mathtime
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

% Text
\setlength{\parindent}{0.5in}
\frenchspacing  \tolerance = 800  \hyphenpenalty = 800

\usepackage{lineno} % Line numbers
\def\linenumberfont{\normalfont\footnotesize\ttfamily}
\setlength\linenumbersep{0.2 in}

\usepackage{setspace}

% Format section and subsection headers
\makeatletter
\renewcommand{\section}{\@startsection
{section}%                   % the name
{1}%                         % the level
{0mm}%                       % the indent
{-\baselineskip}%            % the before skip
{0.5\baselineskip}%          % the after skip
{\normalfont\bf\large}} % the style

\renewcommand{\subsection}{\@startsection
{subsection}%                   % the name
{2}%                         % the level
{0mm}%                       % the indent
{-\baselineskip}%            % the before skip
{0.5\baselineskip}%          % the after skip
{\normalfont\bf}} % the style
\makeatother

% Other
\usepackage{graphicx}
\usepackage[singlelinecheck=false,font=small,labelfont=bf]{caption}
\usepackage[justification=centering]{caption}

% Bibliography
\usepackage[numbers, compress]{natbib} % Bibliography - APA
\bibpunct{(}{)}{;}{a}{}{,}

% Format the Bibliography appropriately
% increase \bibhang to take care of the numbers
\setlength{\bibhang}{0pc}
\makeatletter
%patch \@lbibitem to print the current number before the authors
\patchcmd{\@lbibitem}
  {]}
  {] \theNAT@ctr. \newline }
  {}{}
\makeatother

%%%%%% FRONT MATTER %%%%%%%%%

\title{Comparing methods for estimating parasite-induced host mortality from distributional data: implementation and limitations}
\author{Mark Wilber, Sara Weinstein, and Cherie Briggs}

\begin{document}

\maketitle

\section*{Abstract}

TODO

\doublespacing

\linenumbers
\section*{Introduction}

Infectious agents can have major impacts on animal populations through changing
population dynamics and stability, altering predator-prey interactions, and
even causing species’ decline and extinction. Accurately estimating the impact
of these infectious agents in wildlife is critical to both understanding what
regulates host and parasite populations and making predictions about disease
transmission. The impact of pathogens, such as rabies(), bovine TB(), and
rinderpest(),  is typically quantified based on the presence or absence of
disease, and does not account for the number of infectious agents present.
Although sufficient for many bacterial and viral agents that reproduce within a
host, for macroparasites, hosts cannot be simply categorized as infected and
uninfected because pathology is linked to the intensity of infection (Anderson
and May 1978).  Helminths exhibiting this intensity dependent pathology have
significant impacts on human health (), domestic livestock economics (),
wildlife survival ().  And, while it is generally assumed that some fraction of
wild host populations must succumb to parasitic infections, it is notoriously
difficult to actually quantify parasite-induced host mortality (PIHM) in wild
animal populations. [more]

Ideally, parasite-induced host mortality would be
quantified by experimentally infecting and tracking individual hosts in the
wild population; however, for logistical and ethical reasons this method is
rarely feasible. Data on parasite intensity is much easier to collect and has
often been used to identify the presence PIHM (cite studies) and quantify the
relationship between infection intensity and host mortality (cite studies).

Crofton (1971) first proposed that PIHM could be identified by comparing the
observed parasite distribution in the host population to the distribution
predicted in the absence of parasite-induced host mortality. This method
(“Crofton Method”) assumes that, prior to host mortality, parasites are
distributed in the host population following a negative binomial distribution;
however, as intensity dependent pathology removes heavily infected hosts from
the population, the tail of the distribution is truncated. Mortality is assumed
to not occur in hosts with low intensity infections, thus by iteratively
fitting a negative binomial distribution to hosts with lower and lower parasite
loads, and comparing the tail end of this predicted distribution to the
observed parasite data, one could determine both whether PIHM was occurring and
the parasite distribution in the host population prior to parasite induced
mortality. We give a thorough description and implementation of the Crofton
Method in \emph{Supplementary Information (SI)} 1 and discuss the validity of
its assumptions in the \emph{Discussion}.

The Crofton Method may be able to detect the presence of PIHM however,
quantifying the relationship between infection intensity and survival
probability is more complicated. \cite{Adjei1986} suggested that these values
could be quantified by using the Crofton Method to first estimate the pre-
mortality parasite distribution and then, using those parameters, calculate the
probability of host survival with increasing parasite load. To do this,
\cite{Adjei1986} modeled host survival as a logistic function and then used a
binomial generalized linear model (GLM) to estimate the logistic parameters.
These methods appeared to provide an estimate the parasite intensity at which
50\% of hosts exhibit PIHM (), as well as the unmeasurable fraction of the
population that was lost (SI for adjei method details and extended discussion
of issues?). However, to implement this method the observed data must be
modified to fit the binomial GLM framework and subjectively binned when mean
infection intensity is high or sample sizes are small.

After 30 years, and despite clear limitations \citep{McCallum2000a}, these
methods (particularly the Crofton Method) are still discussed among
parasitologists and are the primary techniques for examining population level
impacts of parasitism using parasite intensity data. In these methods, PIHM can
only be identified by visually examining plots and, with no clear decision
rule; it can be difficult to determine the significance of PIHM across
different host-parasite systems. The survival function produced by the Adjei
Method offers one solution; however, this method requires manipulating the
original data and has never been thoroughly tested.

Estimating parasite impacts on host population using intensity data should only
be done if methods exist to permit unbiased and accurate estimates of PIHM. In
this study, we first propose a novel likelihood method for calculating PIHM. We
next use simulations to compare this likelihood method with previous Adjei
method to (1) test the ability to detect occurrence of PIHM and (2) estimate
lethal parasite load ($LD_{50}$) and associated survival function.  We then
apply both methods to real datasets previously used in PIHM analyses and
compare results. And, finally, we discuss the limitations of inferring PIHM
from distributional data and whether any method for inferring PIHM has a place
in quantitative parasitology.

% Infectious agents can have major impacts on animal populations
% through changing population dynamics and stability \citep{Dobson1992}, altering
% predator-prey interactions \citep{Joly2004}, and even causing species decline
% and extinction \citep{DeCastro2005a,McCallum2012b}. Accurate estimates of
% parasite-induced host mortality (PIHM) in wild animals are important for
% understanding what regulates both host and parasite populations and to make
% predictions about disease transmission in natural systems. Although a negative
% impact on host fitness is a fundamental component of parasitism
% \citep{Lafferty2002} [more], it is notoriously difficult to quantify PIHM in
% wild animal populations \citep{Lester1984,McCallum2000a}.

% To conclusively identify PIHM in wild animal populations, it is necessary to
% experimentally infect and track host populations; a method that is rarely
% possible in most wild animal systems \citep{McCallum2000a}.  Instead, parasitologists
% are often only able to determine the parasite intensity on some number of sampled hosts.  This snapshot, distributional data is far from the
% ideal type of data for addressing questions regarding PIHM, but this is the type of data on which most questions regarding PIHM are asked \citep{Ferguson2011,Royce1990,Lanciani1989,Lester1984,Lester1977}.  In particular, parasitologists are often interested in asking two questions from this data: 1) Is PIHM occurring in a host-parasite system? and if PIHM is occurring 2) Can the effect of parasite intensity on host survival be quantified? While both macro and
% microparasites have detrimental and possibly fatal effects on hosts, for the
% remainder of this paper we limit our discussion to macroparasites that can be
% discretely counted within a host.

% The first of these PIHM questions was addressed by \cite{Crofton1971a} who developed
% a method to test for the presence of PIHM using the truncation of the negative binomial
% distribution.  In short, the Crofton Method assumes that the distribution of
% parasites across hosts before mortality occurs follows a negative binomial
% distribution \citep{AndersonandMay1978,Shaw1998}.  As heavily infected hosts
% begin to die, the negative binomial distribution is truncated and these hosts are no longer observed in a sample. In other words, the
% observed host-parasite distribution and the pre-mortality host-parasite
% distribution will predict substantially different numbers of hosts with high infect intensity (because those have died due to infection), but a similar number of hosts with low infection
% intensity (because those have survived).  \cite{Crofton1971a} noted that by
% starting with all observed hosts and iteratively fitting a negative binomial
% distribution to hosts with lower and lower parasite loads, one could determine
% whether or not PIHM was occurring and estimate the parameters of the host distribution before parasite-induced mortality.  This was done
% graphically by determining whether the parameters of the negative binomial distributions fit to different truncations of the data showed a substantial change as the truncation point moved from heavily infected hosts to lightly infected hosts.  The Crofton Method and this graphical technique are both still currently used to assess whether PIHM is a occurring in a system \citep{Ferguson2011}. We give a thorough description and implementation of the Crofton Method in \emph{Supplementary Information} (SI) 1 and discuss the validity of its assumptions in the \emph{Discussion}.

% The second question regarding PIHM moves beyond a simple yes or no answer and attempts to quantify how parasite intensity affects host survival. \cite{Adjei1986} proposed a method to answer this question
% by first using the Crofton Method to estimate the pre-mortality parameters for
% a host-parasite distribution and then, given these
% parameters, estimating a host survival function that described how the
% probability of host-survival changed with increasing parasite load (see
% \emph{Methods}).  With this host survival function, one can estimate important host-parasite quantities such as the
% parasite intensity at which 50\% of hosts succumb to PIHM ($LD_{50}$), as well
% as the percent of hosts in a population succumbing to PIHM \citep{Adjei1986}.

% Despite these methods both being over three decades old, they are still the
% primary means of answering questions about PIHM given distributional data
% \citep[but see][for an alternative to the Crofton
% Method]{Ferguson2011}. However, both methods have a few important limitations.
% The Crofton Method, and more recent methods \citep{Ferguson2011},
% rely on a visual test to determine whether or not PIHM is occurring in a
% system.  With no clear decision rule, it can be difficult to consistently determine the
% significance of PIHM across different host-parasite systems. In theory, the Adjei Method can ameliorate this problem. In addition to providing information on the host-survival function, this method can also be used to assess the significance of
% PIHM in a system.
% In practice, however, the Adjei Method has never been thoroughly tested and
% relies on a number of questionable data manipulations.

% In this study, we have
% three primary goals. First we wish to test reliability of the Adjei Method for
% answering the aforementioned questions regarding PIHM.  Second, we propose a
% novel method for answering these questions, compare its efficacy against the Adjei Method, and test both methods' ability to detect PIHM on empirical data.  Third, we
% discuss the limitations of inferring PIHM from distributional data alone and
% whether any method for inferring PIHM has a place in the future of
% parasitology.

\section*{Methods}

% \subsection*{The Adjei Method for estimating PIHM}

% The Adjei Method for estimating PIHM has two steps.  The first
% step is to estimate the parameters of the pre-mortality host-parasite
% distribution using the Crofton Method.  The three parameters estimated are the
% total number of hosts before mortality $N_p$,  the mean number of parasites per
% host before mortality $\mu_p$, and the aggregation of parasites before
% mortality given by the parameter $k_p$ from a negative binomial distribution.
% When $k_p$ is small, parasites are highly aggregated among hosts and when
% $k_p$ is large parasites are more evenly spread out \citep{Wilson2002}.  The implementation of the Crofton Method has been discussed at length elsewhere \citep[e.g.][]{Royce1990,Lester1984} and we provide a tested implementation of the method in SI 3.

% The second step of the Adjei Method is to estimate the shape of the host survival function. \cite{Adjei1986} assume that the host survival function follows the logistic form

% \begin{equation}
%     h(x | a, b) = h_x = \dfrac{e^{a - b \log(x)}}{1 + e^{a - b \log(x)}}
%     \label{eq:logistic}
% \end{equation}
% where $x$ is the parasite intensity in a given host and $a$ and $b$ are the two
% parameters of the logistic function. Generally, a larger $a$ allows for hosts to
% tolerate larger parasite intensities before experiencing parasite-induced mortality
% and a larger $b$ leads to a more rapid decline in the probability of host
% survival as parasite intensity increases.  The value $\exp(a / b)$ is referred
% to as the $LD_{50}$, which gives the parasite intensity at which 50\% of hosts
% experience mortality.

% To estimate this function the Adjei Method proceeds as follows.  First, it
% calculates the expected number of hosts with a given parasite load $x$ by using
% the equation $g(x ; \mu_p, k_p) * N_p$, where $g(x ; \cdot)$ is the negative binomial pre-mortality distribution.  Second, the observed and predicted number of hosts
% with $x$ parasites are paired as a single data point and the method then assumes that
% this data point follows a binomial distribution with the total number of
% ``trials'' equal to the predicted number of hosts and the total number of
% ``successes'' equal to the observed number of hosts. In some cases, the
% observed number of hosts is greater than the expected number of hosts and the
% Adjei Method alters the data so that the observed is equal to the predicted
% \citep{Adjei1986}.  After this questionable manipulation, the (observed, predicted) pairs are fit to a standard Generalized Linear Model \citep{McCullagh1989} with a binomial response variable and a logistic link function given by equation \ref{eq:logistic}.  This model provides estimates for parameters $a$, $b$ and $LD_{50}$.

% While not included in the original implementation of the Adjei Method, a
% $\chi^2$ test with a degrees of freedom of 1 can be used to assess whether a GLM model that includes parasite
% intensity as a predictor of host survival probability is a ``better'' model than a
% GLM without this predictor.  This allows the Adjei Method to determine whether
% PIHM is a significant factor in a host-parasite system.

% The Adjei Method's most glaring deficiency is the need to alter the observed
% data in order to fit the model into the binomial GLM framework.  A second more
% subtle problem with the Adjei Method is the potential need to bin data in order
% to predict greater than one host in a given parasite intensity class.  For
% example, if the total number of hosts pre-mortality was 50, the mean number of
% parasites per host pre-mortality was 100 and the aggregation parameter was 1, applying the equation $g(x ; \mu_p=100, k_p=1) * 50$ would result in
% less than 1 individual in all parasite intensities $x$. In other words, the
% Adjei Method cannot be applied to samples with either very high mean parasite
% loads, small sample sizes, or both without some sort of binning of the data.
% While this is not a flaw \emph{per se}, it does add a certain level of
% subjectivity (i.e. which bins should you use?) to a method that already has
% serious potential issues.  In this analysis, we always assume the Adjei Method is not binning the data, though we provide code for applying the binning method in SI 3.

\subsection*{A novel, likelihood method for estimating PIHM}

Here we propose an alternative approach using a likelihood method that does not
require binning or data alteration, potentially reduces the number of
parameters to be estimated, and allows the significance of PIHM to be estimated
using standard statistical techniques.

As with all previously proposed methods for estimating PIHM, the
Likelihood Method first assumes that the pre-mortality distribution follows a
negative binomial distribution $g(x; \mu_p, k_p)$.  The second
assumption is that the host survival function
takes the form of a logistic curve given by
\begin{equation}
    h(x | a, b) = h_x = \dfrac{e^{a - b \log(x)}}{1 + e^{a - b \log(x)}}
    \label{eq:logistic}
\end{equation}
With these two explicit assumptions, the Likelihood Method estimates the 4 parameters $\mu_p$, $k_p$, $a$, and $b$.

To estimate these parameters, we first define a probability distribution that gives the probability of having a parasite load of $x$ parasites conditional on host survival.  Using standard rules of conditional probability this distribution can be written as

\begin{equation}
    P(x | \text{survival}) = \dfrac{P(\text{survival} | x) * P(x)}{P(\text{survival})}
    \label{eq:concept}
\end{equation}

One can see that $P(\text{survival} | x)$ is the survival function $h(x; a, b)$, $P(x)$ is the pre-mortality parasite distribution $g(x; \mu_p, k_p)$ and $P(\text{survival}) = \sum_{x=0}^{\infty} P(\text{survival} | x) * P(x) =  \sum_{x=0}^{\infty} h(x; a, b)  * g(x; \mu_p, k_p)$. Therefore equation \ref{eq:concept} can be written as

\begin{equation}
    P(x | \text{survival}) = \dfrac{h(x; a, b)  * g(x; \mu_p, k_p)}{\sum_{x=0}^{\infty} h(x; a, b)  * g(x; \mu_p, k_p)}
    \label{eq:dist}
\end{equation}

Using this probability distribution, one can then find the parameters $\mu_p$, $k_p$, $a$, and $b$ that maximize the likelihood of an observed host-parasite dataset. Alternatively, one could apply the Crofton Method to estimate $\mu_p$ and $k_p$ and then find the maximum likelihood estimates of $a$ and $b$ and the corresponding $LD_{50}$.

To estimate that significance of PIHM in a host-parasite system, a
likelihood ratio test can be used in which the full model is given by equation
\ref{eq:dist} and the reduced model is given by a negative binomial
distribution.  If PIHM is not significant in the system, the resulting likelihood
ratio statistic should approximately follow a $\chi^2$ distribution with degrees of freedom equal to 2.  We provide the code for implementing this Likelihood Method in SI 3.

\subsection{Evaluating the likelihood and Adjei Methods}

\emph{Question 1: Can we detect PIHM?}

To test the ability of the Adjei Method and the Likelihood Method to identify whether or not PIHM was occurring in a system, we randomly generated data using the following procedure.  First, we drew $N_p$ randomly infected hosts from a
negative binomial distribution with parameters $\mu_p$ and $k_p$.  This represented the dataset observed before mortality. Second, we chose values of $a$ and $b$ and calculated the probability of survival
for all $N_p$ hosts using equation \ref{eq:logistic}.  Third, we drew $N_p$ random numbers from a uniform distribution
between 0 and 1 and if host survival probability was less than this random
number, the host experienced parasite-induced mortality.  The surviving
hosts comprised the dataset that would be obtained in the field, after PIHM.

Using both the pre-mortality and post-mortality simulated datasets,  we assumed
that the values of $N_p$, $\mu_p$, and $k_p$ were known and tested the ability of both methods to correctly determine whether or not PIHM was occurring.  While this scenario is unrealistic because the parameters $N_p$,
$\mu_p$, and $k_p$ are always unknown, we implemented this scenario as a baseline to
establish the efficacy of the methods independent of the estimates of $N_p$, $\mu_p$ and $k_p$.  For the Adjei Method, $N_p$, $\mu_p$, and $k_p$ are estimated using the Crofton Method, while $\mu_p$ and $k_p$ in the Likelihood Method can be estimated jointly with $a$ and $b$ or via the Crofton Method.   If a
method could not correctly predict whether or not PIHM was occurring under these idealistic conditions, we considered this strong evidence of the unreliability of this method.

We used three different values of $\mu_p$ (10, 50, 100) and for each $\mu_p$ we examined three different survival functions that had graduate, moderate, and steep decreases in host survival with increasing parasite intensity (Figure \ref{fig:surv}).  For a given $\mu_p$, each survival function had the same $LD_{50}$ ([$\mu_p = 10$, $LD_{50} = 7.39$], [$\mu_p = 50$, $LD_{50} = 35.57$], [$\mu_p = 100$, $LD_{50}= 77.3$]),  but different values of $a$ and $b$.  We examined each $\mu_p$-survival function pair at  three levels of parasite
aggregation, $k_p = 0.1$, 0.5, and 1 --- realistic values of parasite aggregation in natural populations \citep{Shaw1998}.  For each of these parameter
combinations we simulated 150 datasets and tested the probability of each method incorrectly identifying PIHM in the pre-mortality dataset (Type I error) and correctly identifying PIHM in the post-mortality dataset (power).  For each method, we used a likelihood ratio test to determine whether the full model with PIHM provided a significantly better fit than the reduced model without PIHM at significance level $\alpha = 0.05$.  We tested each parameter combinations for pre-mortality population sizes of $N_p$ = [50, 100, 200, 300, 400, 500]. $N_p$ is not technically the sample size on which the methods are being
tested on the post-mortality data because PIHM reduces $N_p$ for each simulated
dataset.  We therefore used the average number of surviving hosts over all 150 simulations for a given parameter combination as our measure of sample size in the power simulations.

\subsection*{Question 2: Can we estimate fatal parasite intensity and the host survival function?}

To compare the ability of the Adjei Method and the Likelihood Method to recover the $LD_{50}$ and the parameters $a$ and $b$ of the survival function, we used the same simulation procedure and parameter combinations described above. For each parameter
combination we simulated 150 datasets, estimated $a$, $b$, and $LD_{50}$ and calculated the standardized bias and
precision \citep{Walther2005} for these estimates over varying pre-mortality host population sizes  $N_p$ = [300, 500, 1000, 2000, 5000, 7500,
10000]. We used the average number of surviving hosts over all 150 simulations for a given parameter combination as our measure of sample size.  Because parameters $a$ and $b$ showed similar patterns of bias and precision, we only show the results for $a$.

\subsection*{Application to real data}

We tested the ability of the Adjei Method and the Likelihood Method to identify
PIHM in 6 host-parasite datasets given in \cite{Crofton1971a} and 4 datasets
given in \cite{Adjei1986} (Table \ref{table:pihm}). In the \cite{Crofton1971a} datasets, the host was
the snail \emph{Gammarus pulex} which acts as the intermediate host for the
acanthocephalan \emph{Polmorphus minutus}. In the \cite{Adjei1986} datasets,
the hosts were two species of lizard fish \emph{Saurida tumbil} and
\emph{Saurida undosquamis} that were infected by the cestode
\emph{Callitetrarhynchus gracilis}.  Males and females of both fish species
were considered separately.

In both studies, the authors reported PIHM in some of the datasets and we test whether the Adjei
Method and/or the Likelihood Method also predicted PIHM. For the 6 datasets from
\cite{Crofton1971a}, we truncated the data at 4 parasites, applied the Crofton
Method to estimate the pre-mortality distribution, and then ran the Likelihood
Method and Adjei Method using these pre-mortality parameters.  For the
\cite{Adjei1986} datasets, we followed the same procedure as the authors and
first truncated the data at 2 parasites and then fit the Crofton Method for the
female fish of both species.  Then, following the original authors' methods, we parameterized the male pre-mortality
distributions for each species with the results from the females.  Finally, we
applied the Adjei Method and the Likelihood Method to determine whether or not
PIHM was significant for these species and compared our results to those given by the authors.  All fitting to data was done with the Python code provided in SI 3.

\section*{Results}

\subsection*{Question 1: Detecting presence of PIHM}

The Adjei Method showed highly inflated Type I error rates (i.e. falsely detected
PIHM) for all parameter combinations that we
considered (Figure \ref{fig:typeI10}; SI2 Figs 1, 2).  This method also showed the unintuitive pattern of Type I error
rate decreasing as sample size decreased.  This pattern is due to the issue of
binning discussed in the \emph{Introduction} and \emph{SI X}. For small samples sizes, the
applicability of the Adjei Method is compromised without binning the observed
data in some way.  In contrast, the Likelihood Method showed a Type I
error rate at or near the pre-set level of 0.05 for all parameter combinations
and sample sizes considered.

The ability of the Adjei Method to detect PIHM given that it was occurring in a
system (i.e. the power) was close to 1 for larger sample sizes and tended to
decrease as sample size decreased (Figure \ref{fig:typeI10}; SI2 Figs 1, 2).  The Likelihood Method had a power close to
one for all parameter combinations and sample sizes considered.  With gradual
survival functions, the power of the Likelihood Method decreased slightly for small samples sizes (solid black lines; Figure \ref{fig:typeI10}, SI2 Figs 1, 2).

% TODO: Need to figure out what to do with the mu = 100 that are being dropped
% because the sample size is too small!

\subsection*{Question 2: Estimating the $LD_{50}$ and survival function}

The Likelihood Method gave asymptotically unbiased estimates of the $LD_{50}$
for all combinations of parameters examined in this study (Figure \ref{fig:biasld50}, SI2 Fig 3, 4).  Even for
small sample sizes (< 500 hosts), the Likelihood Method's estimate of $LD_{50}$
was largely unbiased, with small biases occurring for host survival functions
that were gradual. The precision of the Likelihood Method's $LD_{50}$ estimates decreased
(increasing coefficient of variation) as sample size decreased for all
parameter combinations we examined (Figure \ref{fig:biasld50}, SI2 Fig 3, 4).

The Adjei Method
always produced biased estimates of the $LD_{50}$ across all parameter combinations (Figure \ref{fig:biasld50}, SI2 Fig 3, 4).  For $\mu_p = 10$, the $LD_{50}$
estimates from the Adjei Method were largely unbiased for large samples sizes,
but showed increasing bias as sample size decreased, particularly for steep
survival functions.  As $\mu_p$ increased, the Adjei method
produced biased estimates of $LD_{50}$ across all sample sizes, with bias
increasing as sample size decreased (SI2 Fig 3, 4). The $LD_{50}$ estimates from the Adjei
Method showed large decreases in precision occurring
for the steepest survival function across all values of $\mu_p$ (Figure \ref{fig:biasld50}, SI2 Fig 3, 4).

In terms of the host survival function, the Likelihood Method gave
asymptotically unbiased estimates of $a$ as sample size increased for
all parameter combinations considered (Figure \ref{fig:biasa}, SI2 Fig 5, 6).  However, as sample size decreased, the
Likelihood Method tended to produce severely biased estimates of $a$.
This was generally more pronounced for steeper survival functions and more
aggregated pre-mortality distributions (e.g. Figure \ref{fig:biasa}).  The Adjei Method produced
biased estimates of $a$ across all sample sizes, with the bias
consistently being larger when the survival function was steeper. The bias of
the Adjei Method's estimate of $a$ also increased as $\mu_p$ increased (Figure \ref{fig:biasa}, SI2 Fig 5, 6).

\subsection*{Application to real data}

Of the 10 datasets we considered, the previous authors visually detected PIHM
in 7 of them (Table \ref{table:pihm}).  The Likelihood Method parameterized
from the pre-mortality parameters of the Crofton Method detected significant
PIHM in 6 of these 7 datasets at a significance level of 0.05.  The only
dataset in which the Likelihood Method did not detect a significant effect of PIHM was the Adjei dataset
for female \emph{S. tumbil}.  For this dataset there was a marginally significant effect
of PIHM ($\chi^2_{df=2} = 5.34; p = 0.069$).

The Adjei Method detected PIHM in 9 of the 10 datasets (Table \ref{table:pihm}),  showing that the Adjei Method has a very high Type I error rate.

\section*{Discussion}

% Sara W Outline
% 5 paragraph disc Outline:
%  Describe benefits of new method (LM)
% No data manipulation
% Estimates fewer parameter
% Statistical output
%  A new method to quantitatively explore mortality using intensity data

% Compare  LM with AM
% LM always does better,
% Detecting  PIHM: more power, less false detection
% Ld50/survival curve: LM more precise/less biased. But these do get worse with decreasing sample size.
% real data , consistent with original author’s results, supports AM type 1 error issues
% AM is fundamentally flawed, LM is clearly better under ideal simulations, but can it be implemented in the real world?

% 3. Data set limitations (for the LM, because no one should even try the AM):
% a. requires large sample sizes (n>X) which may be impossible to collect from the species we would be most interested in getting this data for, eg, large mammals of conservation concern. While 50-100 is very realistic for invertebrates or fish, not so for larger verts
% b. highly aggregated or high mean intensity increase needed sample size (is this right for LM?) which may make this difficult for many helminths with worm burdens in the 100s-1000s
% c. low parasite induced mortality may be hard to detect- particularly an issue for worms in definitive hosts
% d. shape of survival curve matters, although often assumed to be linear (eg Anderson may model) can’t know apriori, and method fails  for both perfectly linear and exceptionally steep survival functions
% e. math works, but not for all parasite-host systems

% 4.even when the math works, should we believe it?
% a. All methods rely on same fundamental assumptions of the crofton method.
% b. assume that pre mort psite dist follows a negative binomial. And this is valid because the NB is extremely flexible and lots of empirical and theoretical work says so.
% c. but, detecting mortality requires that the observed data be significantly different than the NB but because NB is so flexible it can be fit to almost anything, making it impossible to detect pihm
% d.assume truncation from pihm, but could be caused by lots of other things, so detecting significant pihm does not mean it occurred.
% f. serious fundamental issues that need to be recognized and acknowledge with any use of this technique

% 5.(conclusion)Does any of this have a place in quantitative parasitology?

In this study we propose a novel method, the Likelihood Method, for detecting
and describing PIHM from parasite intensity data.  Unlike currently proposed
methods for estimating PIHM, the Likelihood Method does not require any
unnecessary data manipulation, uses fewer parameters [it can, but not always],
and provides a quantitative decision rule for determining whether or not PIHM
is occurring in a host-parasite system.

Using simulated datasets, we show that the Likelihood Method correctly detected
PIHM when it was occurring and correctly failed to detect PIHM when it was not
occurring for sample sizes below 100 hosts. When applied to real data, the
Likelihood Method gave results consistent with what was previously reported. In
contrast, the Adjei Method, the only currently proposed method to estimate the
host survival function and the $LD_{50}$ from parasite intensity data, consistently
detected PIHM even when it was not present and behaved erratically for small sample sizes (< 200
hosts); a consequence of the need to
subjectively bin the data to predict parasite intensity classes with
at least one host.  Finally, when applied to real data, the Adjei Method overestimated the occurrence of PIHM relative to what was previously reported.

In addition to accurately detecting PIHM, the Likelihood Method also provided
unbiased and precise estimates of a host survival function's $LD_{50}$ across
all sample sizes we considered.  While the Likelihood Method did produce
unbiased estimates of the host survival function ($a$ and $b$) for
unrealistically large sample sizes, it gave severely biased estimates of the
host survival function for realistic samples sizes. The bias was most severe
for steep host survival functions due to large changes in the values of $a$ and
$b$ only slightly changing an already steep survival function. The Adjei
Method, on the other hand, gave biased predictions of the $LD_{50}$ and host
survival function across all parameter combinations, with bias increasing with
increasing parasite aggregation and increasing mean parasite intensity.  Taken
together, these results show that Adjei Method is fundamentally flawed and the
Likelihood Method may offer a potential alternative for estimating PIHM.

While we have improved upon the previously existing methods for answering
questions about PIHM, we cannot belie the fact that estimating PIHM from
observational data alone is ladened with assumptions and difficulties
\citep{McCallum2000a}. The most fundamental assumption of all methods for
estimating PIHM is that the shape of the pre-mortality host-parasite
distribution is known and follows a negative binomial distribution. While there
is substantial empirical and theoretical evidence to justify the use of the
negative binomial distribution as the pre-mortality distribution for
macroparasites across hosts \citep{Calabrese2011,Anderson1982a,Shaw1998}, it is
widely recognized that different processes can lead to a variety of
distributions of parasites across hosts
\citep{Isham1995,Grenfell1995b,Wilson2002, Duerr2003}.  However, the critical
assumption of the pre-mortality distribution is not that the processes leading
to the pre-mortality distribution generate a negative binomial distribution,
but rather that the pre-mortality distribution is well-fit by a negative
binomial distribution. The extreme flexibility of the negative binomial
distribution makes it a reasonable candidate distribution for the pre-mortality
distributions.  Therefore, we do not see this assumption as central problem in
any of the proposed methods.

However, to use the pre-mortality distribution to infer whether or not the PIHM
is occurring in a system requires an explicit assumption about the host
survival function and the shape of the post-mortality distribution.  Regarding
the host-survival function, all currently proposed methods of PIHM assume that
the host-survival function is such that uninfected individuals and individuals
with low parasite intensity experience essentially no PIHM. \cite{Lanciani1989}
illustrated the importance of this assumption by showing that when hosts
experienced a linear decrease in survival probability the Crofton Method could
not detect PIHM.   As the most fundamental models of host-parasite dynamics
assume a linear decrease in host survival probability with increasing parasite
intensity \citep{AndersonandMay1978}, the failure of these methods to detect
this relationship is a significant disconnect between empirical and theoretical
disease ecology.  However, empirical work has shown that non-linear functions
of host survival are not uncommon in host-parasite systems \citep{Benesh2011},
so this assumption alone does not preclude the use of PIHM methods on empirical data.

Regarding the shape of the post-mortality distribution, all of these methods
require that the post-mortality distribution be significantly different from a
negative binomial distribution.  This is necessary because none of the above
methods will be able to detect PIHM if a negative binomial distribution is an
adequate fit to the post-mortality distribution.  This is simply because there
will be no need for a more complex model (either truncation of the negative
binomial or the model given in equation \ref{eq:dist}) if a negative binomial
distribution already fits the data. As many observed host-parasite
distributions are not significantly different from a negative binomial
distribution, there may be limited cases where these PIHM methods can even be
considered.

Finally, all of these methods assume that the truncation of a negative binomial
distribution is due to PIHM, but previous studies have shown that a variety of
other processes can lead to the truncation of a negative binomial distribution
such as within host parasite density-dependence, age- dependent variation in
host resistance and heterogeneous rate of infection
\citep{McCallum2000a,Anderson1982a,Rousset1996}.  Therefore, even detecting
``significant'' PIHM in a dataset does not mean that PIHM is cause of the
truncation.

Given these numerous caveats, is the is there a place in parasitology for methods that
estimate PIHM from distributional data?  We are in agreement with
\cite{Lester1984} that, at the very least, methods for estimating PIHM can
provide preliminary insight into whether or not PIHM is worth further
exploration.  However, we stress that these methods should only be used as an
exploratory tool when assessing the role of PIHM in a system and potential
users should critically evaluate whether they think they have a large enough
sample sizes and an appropriate host survival function/post-mortality distribution for the methods developed
in this paper to be applicable.  Even if they are applicable, inferring PIHM
from distributional data is no substitute for field or laboratory experiments
and/or in depth understanding of the natural history of the host-parasite
system under consideration.

\section*{Acknowledgments}

TODO

\singlespacing
\bibliographystyle{/Users/mqwilber/Dropbox/Documents/Bibformats/ecology_letters.bst}
\bibliography{/Users/mqwilber/Dropbox/Documents/Bibfiles/Projects_and_Permits-parasite_host_mortality}

\renewcommand{\arraystretch}{1.2}
\begin{sidewaystable}

    \caption{Comparison of the PIHM predictions of previously used host-parasite datasets to those given by the Adjei Method and the Likelihood Method. The first column specifies the identity of the dataset, the second column specifies whether or not the author indicated that PIHM was occurring in the system, the third column indicates whether or not the Likelihood Method with parameters from the Crofton Method detects significant PIHM, and the final column indicates whether the Adjei Method with pre-mortality parameters estimated from the Crofton Method detects PIHM.  If a method detected significant PIHM the predicted $LD_{50}$ is given in parentheses}

    \centering
    \begin{tabular}{l  p{3cm} p{3cm} l}

    \hline\hline
    Data Set (sample size) & Author detected PIHM? & Likelihood Method?  & Adjei Method? \\

    \hline
    Crofton, Station 1 ($n=538$) & Yes & Yes (7.27) & Yes (9.33) \\
    Crofton, Station 2 ($n=507$) & Yes & Yes (6.92) &  Yes (14.95)\\
    Crofton, Station 3 ($n=633$) & Yes & Yes (5.93) &  Yes (5.98) \\
    Crofton, Station 4 ($n=486$) & No & No &  Yes (7.99) \\
    Crofton, Station 5 ($n=276$) & No & No & Yes (10.58) \\
    Crofton, Station 6 ($n=191$) & No & No & No \\
    Adjei, \emph{S. tumbil} female ($n=446$) & Yes (5.7) & No & Yes (6.37) \\
    Adjei, \emph{S. tumbil} male ($n=452$) & Yes (3.4) & Yes (3.42) & Yes (3.66)  \\
    Adjei, \emph{S. undosquamis} female ($n=2573$) & Yes (3.2) & Yes (3.04) & Yes (3.11) \\
    Adjei, \emph{S. undosquamis} male ($n=2440$) & Yes (1.8) & Yes (1.83) & Yes (1.78) \\

    \end{tabular}
    \label{table:pihm}

\end{sidewaystable}

\begin{figure}
    \includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/conceptual_fig.pdf}

    \caption{Five potential shapes for a host-survival functions. PIHM should be easier to detect for steeper host survival functions \citep{Lanciani1989}, but we may expect the bias in the parameter estimates to increase as it becomes increasingly difficult to distinguish between steep survival functions.}
    \label{fig:surv}
\end{figure}

\begin{figure}

    \includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/typeIpower_figure_for_mu10}

    \caption{The Type I error rate and the power of the Likelihood Method (solid line) and the Adjei Method (dashed lines) when $\mu_p = 10$ for various shapes of the host survival function and levels of aggregation $k_p$.  The first column gives the Type I error rate of each method for falsely detecting PIHM when none is present.  The red line gives the the pre-set Type I error rate of $\alpha = 0.05$.  The second column gives the power of a given method to detect PIHM when it is actually occurring. }
    \label{fig:typeI10}

\end{figure}

\begin{figure}

    \includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/bais_prec_figure_for_ld50_mu10}

    \caption{The bias and the precision of the Likelihood Method (solid line) and the Adjei Method (dashed lines) when $\mu_p = 10$ for various shapes of the host survival function and levels of aggregation $k_p$ when estimating $LD_{50}$.  The first column gives the bias of each method's $LD_{50}$ estimate over 150 simulations. The second column gives the precision of each method's $LD_{50}$ estimate over 150 simulations.}

    \label{fig:biasld50}

\end{figure}

\begin{figure}

    \includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/bais_prec_figure_for_a_mu10}

    \caption{The bias and the precision of the Likelihood Method (solid line) and the Adjei Method (dashed lines) when $\mu_p = 10$ for various shapes of the host survival function and levels of aggregation $k_p$ when estimating the $a$ parameter of the host survival function.  The first column gives the bias of each method's $a$ estimate over 150 simulations. The second column gives the precision of each method's $a$ estimate over 150 simulations.}

    \label{fig:biasa}

\end{figure}

\end{document}