### 2015-08-04 docs/response_to_reviewers.md
Beginning to respond to reviewers

## Response to Reviewers

### Reviewer 1

We thank Reviewer 1 for the helpful suggestions. We agree with Reviewer 1 that host-parasite distributions are dynamic and that it would be interesting to apply these models to

### Reviewer 2

We thank Reviewer 2 for their helpful feedback and address each of their points in turn.

1. The major critique provided by Reviewer 2 was that we were too negative about our results in the Discussion and the Abstract. We agree with this comment and have reworded the Abstract and the Discussion to reflect why we think .  That being said, we do still feel it is important to highlight the limitations of